---
title: "ANSC516_FinalProject"
author: "Sana Mohammad"
date: "2025-04-22"
output: html_document
---

# ANSC516- Final Project

The following R Markdown contains the code used in conjunction with qiime2 outputs from my slurm files to create figures used in my final project.

# Setup

Clean the environment and set working directory if needed.

```{r}
rm(list = ls ())
getwd()
```

## Packages

```{r, message=F}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#install.packages("BiocManager")
#BiocManager::install("phyloseq")

library(tidyverse)
library(qiime2R)
library(ggpubr)
library(readxl)
library(dplyr)
library(devtools)
library(vegan)
```

## Loading Data

### Metadata

```{r}
meta <- read_excel("SraRunTable.xlsx")
meta <- meta %>%
  select(id, Host, isolation_source)
head(meta)
```

Renaming factors in metadata

```{r}
meta <- meta %>% 
  mutate(Host = case_when(
  Host == "not applicable" ~    "na",
  Host == "Solanum lycopersicum cv. Hawaii7996 x S. pimpinellifolium West Virginia700 RIL line" ~    "HA7996 x WV700 RIL line",
  Host == "S. lycopersicum cv. Hawaii7996 x S. pimpinellifolium West Virginia700 RIL line" ~    "HA7996 x WV700 RIL line",
  Host == "S. lycopersicum cv. Pearson" ~    "Pearson",
  Host == "Never-ripe (Nr) mutant in S. lycopersicum cv. Pearson background" ~    "Nr (Pearson)",
  Host == "NahG transgenic mutant in S. lycopersicum cv. Money Maker background" ~    "NahG (MM)",
  Host == "ACD transgenic mutant in S. lycopersicum cv. UC82B background" ~    "ACD (UC82B)",
  Host == "S. lycopersicum cv. Bonnie Best" ~    "Bonnie Best",
  Host == "S. lycopersicum cv. Castle Mart II" ~    "CastleMart",
  Host == "Bulk soil from hormone mutant pots" ~    "Bulk Soil",
  Host == "Bulk soil from RIL pots" ~    "Bulk Soil",
  Host == "defenseless-1 (def1) mutant in S. lycopersicum cv. Castle Mart II background" ~    "def1 (CM)",
  Host == "S. lycopersicum cv. Hawaii7996" ~    "Hawaii7996",
  Host == "S. lycopersicum cv. Money Maker" ~    "Money Maker",
  Host == "S. lycopersicum cv. UC82B" ~    "UC82B",
  Host == "S. pimpinellifolium accession LA2093" ~    "LA2093",
  Host == "S. pimpinellifolium West Virginia700" ~    "West Virginia700",
  TRUE                      ~ "other"
)) %>%
  mutate(isolation_source = case_when(
    isolation_source == "bulk soil from pots of natural soil mixed with potting mix\\, in greenhouse" ~ "Bulk Soil",
    isolation_source == "root endosphere of greenhouse grown plants in natural soil mixed with potting mix" ~ "Root Endosphere",
    isolation_source == "root rhizosphere of greenhouse grown plants in natural soil mixed with potting mix" ~ "Root Rhizosphere",
    isolation_source == "not applicable" ~ "na",
  ))

head(meta)
```

### Alpha Diversity vectors

```{r}
evenness = read_qza("core-metrics-results/evenness_vector.qza")
evenness <-evenness$data %>% rownames_to_column("id")

observed_features = read_qza("core-metrics-results/observed_features_vector.qza")
observed_features<-observed_features$data %>% rownames_to_column("id")

shannon = read_qza("core-metrics-results/shannon_vector.qza")
shannon<-shannon$data %>% rownames_to_column("id")

faith_pd = read_qza("core-metrics-results/faith_pd_vector.qza")
faith_pd<-faith_pd$data %>% rownames_to_column("id")

#simpson = read_qza("core-metrics-results/simpson_vector.qza")
#simpson<-simpson$data %>% rownames_to_column("id")

chao = read_qza("core-metrics-results/chao1_vector.qza")
chao<-chao$data %>% rownames_to_column("id")
```

### Beta Diversity Results

```{r}
bc_PCoA<-read_qza("core-metrics-results/bray_curtis_pcoa_results.qza")
wUF <- read_qza("core-metrics-results/weighted_unifrac_pcoa_results.qza")
UF <- read_qza("core-metrics-results/unweighted_unifrac_pcoa_results.qza")
jacc <- read_qza("core-metrics-results/jaccard_pcoa_results.qza")
```

# Alpha Diversity

Merging alpha diversity results with the metadata

```{r}
observed_features$observed_features <- as.numeric(observed_features$observed_features)

str(observed_features)
```

```{r}
alpha_diversity = merge(x=faith_pd, y=evenness, by.x = "id", by.y = "id")
alpha_diversity = merge(alpha_diversity, observed_features, by.x = "id", by.y = "id")
alpha_diversity = merge(alpha_diversity, shannon, by.x = "id", by.y = "id")

meta = merge(meta, alpha_diversity, by.x = "id", by.y = "id")
row.names(meta) <- meta$id

head(meta)
```

## Normality Testing

### Shannon

```{r}
ggqqplot(alpha_diversity$shannon_entropy)
shapiro.test(alpha_diversity$shannon_entropy)
```

Not normal

### Faith PD

```{r}
ggqqplot(alpha_diversity$faith_pd)
shapiro.test(alpha_diversity$faith_pd)
```

Not normal

### Evenness

```{r}
ggqqplot(alpha_diversity$pielou_evenness)
shapiro.test(alpha_diversity$pielou_evenness)
```

Not normal

### Observed Features

```{r}
ggqqplot(alpha_diversity$observed_features)
shapiro.test(alpha_diversity$observed_features)
```

Not normal

None of the alpha diversity metrics are normal, so I will use non-parametric statistical tests (Kruskal-Wallis) moving forward.

## Plots

### Shannon

```{r}
plot_shan <- meta %>%
  filter(!Host == "na") %>%
  group_by(Host, isolation_source) %>%
  arrange(shannon_entropy) %>%
ggplot(aes(x=shannon_entropy, y=isolation_source, color=Host))+
  #facet_wrap(~isolation_source, ncol=3) +
  geom_boxplot(position=position_dodge(1))+
  theme_bw()+
  xlab("Shannon Entropy")+
  ylab("")+
  ggtitle("Shannon Index")+
  theme(plot.title=element_text(hjust=0.5),
        aspect.ratio = 1
        #legend.position="none"
        )+
  coord_flip()
  

#ggsave("shannon.tiff")
```

### Faith PD

```{r}
plot_fpd <- meta %>%
  filter(!Host == "na") %>%
  group_by(Host, isolation_source) %>%
  arrange(faith_pd) %>%
ggplot(aes(x=faith_pd, y=isolation_source, color=Host))+
  #facet_wrap(~isolation_source, ncol=3) +
  geom_boxplot(position=position_dodge(1))+
  theme_bw()+
  xlab("Faith PD")+
  ylab("")+
  ggtitle("Faith Phylogenetic Diversity")+
  theme(plot.title=element_text(hjust=0.5),
        aspect.ratio = 1
        #legend.position="none"
        )+
  coord_flip()

#ggsave("faithPD.tiff")
```

### Evenness

```{r}
plot_even <- meta %>%
  filter(!Host == "na") %>%
  group_by(Host, isolation_source) %>%
  arrange(pielou_evenness) %>%
ggplot(aes(x=pielou_evenness, y=isolation_source, color=Host))+
  #facet_wrap(~isolation_source, ncol=3) +
  geom_boxplot(position=position_dodge(1))+
  theme_bw()+
  xlab("Evenness")+
  ylab("")+
  ggtitle("Pielou's Evenness")+
  theme(plot.title=element_text(hjust=0.5),
        aspect.ratio = 1
        #legend.position="none"
        )+
  coord_flip()

#ggsave("evenness.tiff")
```

### Observed Features

```{r}
plot_obs <- meta %>%
  filter(!Host == "na") %>%
  group_by(Host, isolation_source) %>%
  arrange(observed_features) %>%
ggplot(aes(x=observed_features, y=isolation_source, color=Host))+
  #facet_wrap(~isolation_source, ncol=3) +
  geom_boxplot(position=position_dodge(1))+
  theme_bw()+
  xlab("Observed Features")+
  ylab("")+
  ggtitle("Observed Features")+
  theme(plot.title=element_text(hjust=0.5),
        aspect.ratio = 1
        #legend.position="none"
        )+
  coord_flip()

#ggsave("obsFeat.tiff")
```

### Arranging Figure

```{r}
ggarrange(plot_shan, plot_fpd, plot_even, plot_obs, ncol=2, nrow=2, common.legend=T, legend="right")

ggsave("alpha.tiff")
```

## Statistical Analysis

```{r}
endo <- meta %>%
  filter(isolation_source=="Root Endosphere")

rhizo <- meta %>%
  filter(isolation_source=="Root Rhizosphere")
```

### Endosphere

```{r}
kruskal.test(shannon_entropy ~ Host, data=endo)
```

```{r}
kruskal.test(faith_pd ~ Host, data=endo)
```

```{r}
kruskal.test(pielou_evenness ~ Host, data=endo)
```

```{r}
kruskal.test(observed_features ~ Host, data=endo)
```

### Rhizosphere

```{r}
kruskal.test(shannon_entropy ~ Host, data=rhizo)
```

```{r}
kruskal.test(faith_pd ~ Host, data=rhizo)
```

```{r}
kruskal.test(pielou_evenness ~ Host, data=rhizo)
```

```{r}
kruskal.test(observed_features ~ Host, data=rhizo)
```

# Beta Diversity

## Ordination Plots

Merging results with the metadata

```{r}
bc_meta <- bc_PCoA$data$Vectors %>%
  select(SampleID, PC1, PC2, PC3) %>%
  inner_join(meta, by = c("SampleID" = "id")) %>%
  filter(!Host=="na")

jacc_meta <- jacc$data$Vectors %>%   
  select(SampleID, PC1, PC2, PC3) %>%   
  inner_join(meta, by = c("SampleID" = "id")) %>%   
  filter(!Host=="na")

wUF_meta <- wUF$data$Vectors %>%   
  select(SampleID, PC1, PC2, PC3) %>%   
  inner_join(meta, by = c("SampleID" = "id")) %>%   
  filter(!Host=="na")

UF_meta <- UF$data$Vectors %>%   
  select(SampleID, PC1, PC2, PC3) %>%   
  inner_join(meta, by = c("SampleID" = "id")) %>%   
  filter(!Host=="na")
```

### Bray-Curtis

```{r}
brayPlot <- ggplot(bc_meta, aes(x=PC1, y=PC2, 
                                color=Host, shape=isolation_source)) +
  geom_point(size = 2,
             alpha= 1) +
  xlab(paste0("PC1 (", round(100*bc_PCoA$data$ProportionExplained[1], digits = 2), "%)")) +
  ylab(paste0("PC2 (", round(100*bc_PCoA$data$ProportionExplained[2], digits = 2), "%)")) +
  scale_color_discrete(name = "Host")+
  scale_shape_manual(values=c(3,16,17),
                     name="Isolation Source")+
  ggtitle("Bray-Curtis")+
  theme_bw()+
  theme(plot.title = element_text(hjust=0.5))
  
brayPlot
```

#### Jaccard

```{r}
jaccPlot <- ggplot(jacc_meta, aes(x=PC1, y=PC2, 
                                color=Host, shape=isolation_source)) +
  geom_point(size = 2,
             alpha= 1) +
  xlab(paste0("PC1 (", round(100*bc_PCoA$data$ProportionExplained[1], digits = 2), "%)")) +
  ylab(paste0("PC2 (", round(100*bc_PCoA$data$ProportionExplained[2], digits = 2), "%)")) +
  scale_color_discrete(name = "Host")+
  scale_shape_manual(values=c(3,16,17),
                     name="Isolation Source")+
  ggtitle("Jaccard")+
  theme_bw()+
  theme(plot.title = element_text(hjust=0.5))
  
jaccPlot
```

#### Weighted Unifrac

```{r}
weighPlot <- ggplot(wUF_meta, aes(x=PC1, y=PC2, 
                                color=Host, shape=isolation_source)) +
  geom_point(size = 2,
             alpha= 1) +
  xlab(paste0("PC1 (", round(100*bc_PCoA$data$ProportionExplained[1], digits = 2), "%)")) +
  ylab(paste0("PC2 (", round(100*bc_PCoA$data$ProportionExplained[2], digits = 2), "%)")) +
  scale_color_discrete(name = "Host")+
  scale_shape_manual(values=c(3,16,17),
                     name="Isolation Source")+
  ggtitle("Weighted Unifrac")+
  theme_bw()+
  theme(plot.title = element_text(hjust=0.5))
  
weighPlot
```

#### Unweighted Unifrac

```{r}
unweightPlot <- ggplot(UF_meta, aes(x=PC1, y=PC2, 
                                color=Host, shape=isolation_source)) +
  geom_point(size = 2,
             alpha= 1) +
  xlab(paste0("PC1 (", round(100*bc_PCoA$data$ProportionExplained[1], digits = 2), "%)")) +
  ylab(paste0("PC2 (", round(100*bc_PCoA$data$ProportionExplained[2], digits = 2), "%)")) +
  scale_color_discrete(name = "Host")+
  scale_shape_manual(values=c(3,16,17),
                     name="Isolation Source")+
  ggtitle("Unweighted Unifrac")+
  theme_bw()+
  theme(plot.title = element_text(hjust=0.5))
  
unweightPlot
```

#### Arranging Figure

```{r}
ggarrange(brayPlot, jaccPlot, weighPlot, unweightPlot, 
          ncol=2, nrow=2, 
          common.legend=T, legend="right")
```

## Statistics

### Bray-Curtis

Loading distance matrix and merging with metadata

```{r}
bc_dist_mat<-read_qza("core-metrics-results/bray_curtis_distance_matrix.qza")
bc_dm <- as.matrix(bc_dist_mat$data) 
rownames(bc_dm) == meta$id ## all these values need to be "TRUE"
```

```{r}
metadata_sub <- meta[match(rownames(bc_dm),meta$id),]
rownames(bc_dm) == metadata_sub$id ## all these values need to be "TRUE"
```

```{r}
PERMANOVA_bray <- adonis2(bc_dm ~ Host*isolation_source, data = metadata_sub)

write.table(PERMANOVA_bray,"output/bray_Adonis_overall.csv",sep=",", row.names = TRUE) 
```

### Jaccard

```{r}
jacc_dist_mat<-read_qza("core-metrics-results/jaccard_distance_matrix.qza")
jacc_dm <- as.matrix(jacc_dist_mat$data) 
rownames(jacc_dm) == meta$id

metadata_subj <- meta[match(rownames(jacc_dm),meta$id),]
rownames(jacc_dm) == metadata_subj$id

PERMANOVA_jacc <- adonis2(jacc_dm ~ Host*isolation_source, data = metadata_subj)

write.table(PERMANOVA_jacc,"output/jaccard_Adonis_overall.csv",sep=",", row.names = TRUE) 
```

### Weighted Unifrac

```{r}
wu_dist_mat<-read_qza("core-metrics-results/weighted_unifrac_distance_matrix.qza")
wu_dm <- as.matrix(wu_dist_mat$data) 
rownames(wu_dm) == meta$id

metadata_subw <- meta[match(rownames(wu_dm),meta$id),]
rownames(wu_dm) == metadata_subw$id

PERMANOVA_wu <- adonis2(wu_dm ~ Host*isolation_source, data = metadata_subw)

write.table(PERMANOVA_wu,"output/wu_Adonis_overall.csv",sep=",", row.names = TRUE) 
```

### Unweighted Unifrac

```{r}
u_dist_mat<-read_qza("core-metrics-results/unweighted_unifrac_distance_matrix.qza")
u_dm <- as.matrix(u_dist_mat$data) 
rownames(u_dm) == meta$id

metadata_subu <- meta[match(rownames(u_dm),meta$id),]
rownames(u_dm) == metadata_subu$id

PERMANOVA_u <- adonis2(u_dm ~ Host*isolation_source, data = metadata_subu)

write.table(PERMANOVA_u,"output/u_Adonis_overall.csv",sep=",", row.names = TRUE)
```

### Pairwise PERMANOVA

Creating pairwise function

```{r}
pairwise.adonis2 <- function(x, data, strata = NULL, nperm=999, ... ) {
  
  #describe parent call function 
  ststri <- ifelse(is.null(strata),'Null',strata)
  fostri <- as.character(x)
  #list to store results
  
  #copy model formula
  x1 <- x
  # extract left hand side of formula
  lhs <- x1[[2]]
  # extract factors on right hand side of formula 
  rhs <- x1[[3]]
  # create model.frame matrix  
  x1[[2]] <- NULL   
  rhs.frame <- model.frame(x1, data, drop.unused.levels = TRUE) 
  
  # create unique pairwise combination of factors 
  co <- combn(unique(as.character(rhs.frame[,1])),2)
  
  # create names vector   
  nameres <- c('parent_call')
  for (elem in 1:ncol(co)){
    nameres <- c(nameres,paste(co[1,elem],co[2,elem],sep='_vs_'))
  }
  #create results list  
  res <- vector(mode="list", length=length(nameres))
  names(res) <- nameres
  
  #add parent call to res 
  res['parent_call'] <- list(paste(fostri[2],fostri[1],fostri[3],', strata =',ststri, ', permutations',nperm ))
  
  
  #start iteration trough pairwise combination of factors  
  for(elem in 1:ncol(co)){
    
    #reduce model elements  
    if(inherits(eval(lhs),'dist')){	
      xred <- as.dist(as.matrix(eval(lhs))[rhs.frame[,1] %in% c(co[1,elem],co[2,elem]),
                                           rhs.frame[,1] %in% c(co[1,elem],co[2,elem])])
    }else{
      xred <- eval(lhs)[rhs.frame[,1] %in% c(co[1,elem],co[2,elem]),]
    }
    
    mdat1 <-  data[rhs.frame[,1] %in% c(co[1,elem],co[2,elem]),] 
    
    # redefine formula
    if(length(rhs) == 1){
      xnew <- as.formula(paste('xred',as.character(rhs),sep='~'))	
    }else{
      xnew <- as.formula(paste('xred' , 
                               paste(rhs[-1],collapse= as.character(rhs[1])),
                               sep='~'))}
    
    #pass new formula to adonis
    if(is.null(strata)){
      ad <- adonis2(xnew,data=mdat1, ... )
    }else{
      perm <- how(nperm = nperm)
      setBlocks(perm) <- with(mdat1, mdat1[,ststri])
      ad <- adonis2(xnew,data=mdat1,permutations = perm, ... )}
    
    res[nameres[elem+1]] <- list(ad[1:5])
  }
  #names(res) <- names  
  class(res) <- c("pwadstrata", "list")
  return(res)
} 
```

```{r}
bray_pair <- pairwise.adonis2(bc_dm ~ isolation_source, data = metadata_sub)

#write.table(bray_pair,"output/bray_Adonis_pairwise.csv",sep=",",
 #           row.names = T) 

bray_pair
```

```{r}
jacc_pair <- pairwise.adonis2(jacc_dm ~ isolation_source, data = metadata_subj)

jacc_pair
```

```{r}
wu_pair <- pairwise.adonis2(wu_dm ~ isolation_source, data = metadata_subw)

wu_pair
```

```{r}
u_pair <- pairwise.adonis2(u_dm ~ isolation_source, data = metadata_subu)

u_pair
```

# Taxonomy

```{r}
#library(qiime2R)
library(phyloseq)
library(zoo)
#library(tidyverse)
library(glue)
```

```{r}
getwd()
```

Creating a directory for the output.

```{r}
dir.create("output/taxa")
```

Reading in taxonomy.qza file from Qiime2

```{r}
taxonomy<-read_qza("taxonomy.qza")
head(taxonomy$data)
```

```{r}
tax.clean<-parse_taxonomy(taxonomy$data)
head(tax.clean)
```

Making NA's + ASVs that are unclassified at any level classified as the lowest taxonomic level for which there is a classification.

```{r}
tax.clean[is.na(tax.clean)] <- ""
for (i in 1:nrow(tax.clean)){
  if (tax.clean[i,2] == ""){
    kingdom <- paste("Kingdom_", tax.clean[i,1], sep = "")
    tax.clean[i, 2:7] <- kingdom
  } else if (tax.clean[i,3] == ""){
    phylum <- paste("Phylum_", tax.clean[i,2], sep = "")
    tax.clean[i, 3:7] <- phylum
  } else if (tax.clean[i,4] == ""){
    class <- paste("Class_", tax.clean[i,3], sep = "")
    tax.clean[i, 4:7] <- class
  } else if (tax.clean[i,5] == ""){
    order <- paste("Order_", tax.clean[i,4], sep = "")
    tax.clean[i, 5:7] <- order
  } else if (tax.clean[i,6] == ""){
    family <- paste("Family_", tax.clean[i,5], sep = "")
    tax.clean[i, 6:7] <- family
  } else if (tax.clean[i,7] == ""){
    tax.clean$Species[i] <- paste("Genus",tax.clean$Genus[i], sep = "_")
  }
}
```

## Taxa Bar Plot

Importing tree data

```{r, eval=FALSE}
physeq <- qza_to_phyloseq(
  features="core-metrics-results/rarefied_table.qza",
  tree="rooted-tree.qza",
  taxonomy = "taxonomy.qza",
  metadata = "SraRunTable.txt"
)

metadata <-read_excel("SraRunTable.xlsx")
head(metadata)
metadata <- metadata[-1,]

#rownames(metadata) <- metadata[ ,1]

#metadata[,1]

rownames(metadata) <- metadata$id

rownames(metadata)

metadata <- meta
```

```{r eval=F}
physeq_otu_table <- data.frame(otu_table(physeq), check.names = F)

tax.clean = tax.clean[row.names(tax.clean) %in% rownames(physeq_otu_table),]
metadata.filtered = metadata[row.names(metadata) %in% colnames(physeq_otu_table),]

OTU.physeq = otu_table(as.matrix(physeq_otu_table), taxa_are_rows=TRUE)

tax.physeq = tax_table(as.matrix(tax.clean))    
meta.physeq = sample_data(metadata.filtered)

physeq_bar_plot = phyloseq(OTU.physeq, tax.physeq, meta.physeq)

view(OTU.physeq)
view(tax.physeq)
view(meta.physeq)
view(physeq_bar_plot)
```

```{r}
my_colors <- c(
  '#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c',
  '#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#ffff99','#b15928', 
  "#CBD588", "#5F7FC7", "orange","#DA5724", "#508578", "#CD9BCD",
  "#AD6F3B", "#673770","#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F","#D1A33D", "#8A7C64", "#599861", "gray", "black"
)
```

```{r, eval=F}
my_level <- c("Phylum", "Family", "Genus")
my_column <- "isolation_source"
```

```{r,eval=F}
rm(taxa.summary)

abund_filter <- 0.05  # Our abundance threshold
ml ="Genus"


```

```{r, eval=F}
for(ml in my_level){
  print(ml)
  
  taxa.summary <- physeq_bar_plot %>%
    tax_glom(taxrank = ml, NArm = FALSE) %>%  # agglomerate at `ml` level
    transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
    psmelt()  %>%                               # Melt to long format
    group_by(get(my_column), get(ml)) %>%
    summarise(Abundance.average=mean(Abundance)) 
  taxa.summary <- as.data.frame(taxa.summary)
  colnames(taxa.summary)[1] <- my_column
  colnames(taxa.summary)[2] <- ml
  
  physeq.taxa.max <- taxa.summary %>% 
    group_by(get(ml)) %>%
    summarise(overall.max=max(Abundance.average))
  
  physeq.taxa.max <- as.data.frame(physeq.taxa.max)
  colnames(physeq.taxa.max)[1] <- ml
  
  # merging the phyla means with the metadata #
  physeq_meta <- merge(taxa.summary, physeq.taxa.max)
  
  
  physeq_meta_filtered <- filter(physeq_meta, overall.max>abund_filter)
  #str(physeq_meta_filtered)
  
  physeq_meta_filtered$body.site.ord = factor(physeq_meta_filtered$body.site, c("left palm", "right palm", "gut", "tongue"))
  
  # Plot 
  ggplot(physeq_meta_filtered, aes(x = get(my_column), y = Abundance.average, fill = get(ml))) + 
    #facet_grid(.~LitterTreatment) +
    geom_bar(stat = "identity") +
    scale_fill_manual(values = my_colors) +
    # Remove x axis title
    #theme(axis.title.x = element_blank()) + 
    ylim(c(0,1)) +
    guides(fill = guide_legend(reverse = F, keywidth = .5, keyheight = .5, ncol = 1)) +
    theme(legend.text=element_text(size=8)) +
    #theme(legend.position="bottom") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    theme(legend.title = element_blank()) +
    ylab("Relative Abundance") +
    xlab(my_column) +
    ggtitle(paste0(ml, " (>", abund_filter * 100,"%) in at least 1 sample")) 
  ggsave(paste0("output/taxa/", ml, "BarPlot_", my_column, ".png"), height = 5, width = 4)
}

```

# Cooccurance

```{r}
library(Hmisc)
library(plyr)
library(reshape2)
```

```{r}
ASVs <- read_qza("table.qza")
ASV_table <- as.data.frame(ASVs$data)
```

```{r}
ASV_table$ASVnos <- paste0("ASV", 1:nrow(ASV_table))
ASV_table$ASVstring <- rownames(ASV_table)
rownames(ASV_table) <- ASV_table$ASVnos
ASVkey <- ASV_table[, (ncol(ASV_table)-1):ncol(ASV_table)]
ASV_table <- ASV_table[,-(ncol(ASV_table)-1):-ncol(ASV_table)]
```

```{r}
dataset <- as.data.frame(t(ASV_table))
```

```{r}
head(dataset[,1:10])

#metadata<-read_excel("SraRunTable.xlsx")
#metadata <- metadata[-1,]
#str(metadata)
#colnames(metadata)[3] = "Host"
metadata <- meta

dataset <- merge(metadata, dataset, by.x = "id", by.y = 0)
treatments<-as.vector(unique(dataset$Host))
datasetn<-dataset
datasetn[datasetn==0]<-NA
```

```{r}
summary(metadata$isolation_source)
```

```{r}
num_metadata_columns <- ncol(metadata)

q_cutoff <- 0.05

final_results<-data.frame()

my_column <- "isolation_source"
n1 <- 4
n2 <- 4
n3 <- 4
n4 <- 4
```

```{r}
for(i in 1:length(treatments)){
  #subset the data for a particular treatment YOU MUST ENTER THE HEADER OF THE COLUMN THAT HAS THE DIFFERENT TREATMENTS IN THIS CASE “Foaming_Status”
  print(paste("reading ",treatments[i],sep=""))
  temp<-subset(dataset, get(my_column)==treatments[i])
  tempn<-subset(datasetn, get(my_column)==treatments[i])
  print(paste("finished reading ",treatments[i],sep=""))
  # making an object that has all the results in it (both rho and P values)
  results<-rcorr(as.matrix(temp[,-c(1:num_metadata_columns)]),type="spearman") ## use the "-c" parameter to remove metadata columns
  resultsn<-rcorr(as.matrix(tempn[,-c(1:num_metadata_columns)]),type="spearman")
  
  #make two seperate objects for p-value and correlation coefficients
  rhos<-results$r
  ps<-results$P
  ns<-resultsn$n
  # going to melt these objects to 'long form' where the first two columns make up the pairs of OTUs, I am also removing NA's as they are self-comparisons, not enough data, other bad stuff
  ps_melt<-na.omit(melt(ps))
  #creating a qvalue based on FDR
  ps_melt$qval<-p.adjust(ps_melt$value, method = "BH")
  #making column names more relevant
  
  names(ps_melt)[3]<-"pval"
  # if you are of the opinion that it is a good idea to subset your network based on adjusted P-values (qval in this case), you can then subset here
  ps_sub<-subset(ps_melt, qval < q_cutoff)
  
  # now melting the rhos, note the similarity between ps_melt and rhos_melt
  rhos_melt<-na.omit(melt(rhos))
  names(rhos_melt)[3]<-"rho"
  
  # now melting the ns
  ns_melt<-(melt(ns))
  names(ns_melt)[3]<-"n"
  
  #merging together and remove negative rhos
  merged<-merge(ps_sub,rhos_melt,by=c("Var1","Var2"))
  if (treatments[i]==treatments[1]) {
    merged<-merge(merged,subset(ns_melt, n > n1),by=c("Var1","Var2"))
  }   else if (treatments[i]==treatments[2]) {
    merged<-merge(merged,subset(ns_melt, n > n2),by=c("Var1","Var2"))
  }   else if (treatments[i]==treatments[3]) {
    merged<-merge(merged,subset(ns_melt, n > n3),by=c("Var1","Var2"))
  }   else if (treatments[i]==treatments[4]) {
    merged<-merge(merged,subset(ns_melt, n > n4),by=c("Var1","Var2"))
  }   else
    print("Somethings wrong with your treatment designations. Please Check!!")
  
  if (nrow(merged) > 0) {
    merged$trt<-treatments[i]
    final_results<-rbind(final_results, merged)
  }   else {
    print("no correlations for this variable")
  }
  
  print(paste("finished ",treatments[i],sep=""))
}
```

# Taxonomy

```{r}
head(tax.clean)


```
